comgen-outline

### 2017-10-31 paper/appendix-bda.Rmd
move bda details to appendix

---
title             : "The language of generalization"
shorttitle        : "The language of generalization"

author: 
  - name          : "Michael Henry Tessler"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "450 Serra Mall, Bldg. 420, Rm. 316, Stanford, CA 94305"
    email         : "mtessler@stanford.edu"
  - name          : "Noah D. Goodman"
    affiliation   : "1"

affiliation:
  - id            : "1"
    institution   : "Department of Psychology, Stanford University"

header-includes:
  - \usepackage{tabularx}
  - \usepackage{multicol}
  - \usepackage{wrapfig}
  - \usepackage{caption}
  - \usepackage{booktabs}
  - \usepackage{amsmath}
  - \usepackage{graphicx}
  - \usepackage{subcaption}

author_note: >
  This manuscript is currently in prep. Comments or suggestions should be directed to MHT.

abstract: > 
    Generalizations are central to human understanding and language provides simple ways to convey them (e.g., "Dogs bark").
    The language of generalization is ubiquitous in everyday discourse and child-directed speech. 
    Yet the meaning of these expressions is logically puzzling (e.g., not all birds fly) and has resisted precise formalization.
    The major issue in formalizing generalization in language is in determining what which such statements are true or false.
    Using a probabilistic model of pragmatic reasoning, we explore the hypothesis that the meaning of such linguistic expressions is *simple but underspecified*, and that general communicative principles can be used to establish a more precise meaning in context. 
    To test this theory, we examine endorsements of generalizations about three different domains: generalizations about categories (*generic language* e.g., "Dogs bark"), events (*habitual language* e.g., "John runs"), and causes (*causal language* e.g., "Staring at the sun makes you go blind"). 
    Across these diverse domains, we find that our model explains almost all of the variance in human endorsements of generalizations in language, while simpler models fall short.
    This is the first formal theory that makes precise, quantitative predictions about human understanding of generalizations in language.
    The results suggest that central phenomena emerge from the interaction of diverse prior beliefs about properties with general communicative principles.

keywords          : "genericity, generalization, generics, pragmatics, semantics, bayesian modeling"
wordcount         : "20000"

bibliography      : ["generics.bib"]

figsintext        : yes
figurelist        : no
tablelist         : no
footnotelist      : no
lineno            : no

lang              : "english"
class             : "man"
output            : papaja::apa6_pdf
---

\newcommand{\denote}[1]{\mbox{ $[\![ #1 ]\!]$}}
\newcommand*\diff{\mathop{}\!\mathrm{d}}
\definecolor{Red}{RGB}{255,0,0}
\definecolor{Green}{RGB}{10,200,100}
\definecolor{Blue}{RGB}{10,100,200}

\newcommand{\mht}[1]{{\textcolor{Blue}{[mht: #1]}}}
\newcommand{\ndg}[1]{{\textcolor{Green}{[ndg: #1]}}}
\newcommand{\red}[1]{{\textcolor{Red}{#1}}}

```{r global_options, include=FALSE}
rm(list=ls())
knitr::opts_chunk$set(fig.width=6, fig.height=5, fig.crop = F, fig.path='figs/',
                      echo=FALSE, warning=FALSE, cache=T, message=FALSE, sanitize = T)

```

```{r libraries, cache = F}
library(papaja)
library(formatR)
library(rwebppl)
library(xtable)
library(tidyverse)
library(forcats)
library(langcog)
# library(dplyr)
library(data.table)
library(coda)
library(ggthemes)
library(ggrepel)
library(jsonlite)
library(gridExtra)
library(lme4)
library(knitr)
library(cowplot)
theme_set(theme_few())
estimate_mode <- function(s) {
  d <- density(s)
  return(d$x[which.max(d$y)])
}
hdi_upper<- function(s){
  m <- HPDinterval(mcmc(s))
  return(m["var1","upper"])
}
hdi_lower<- function(s){
  m <- HPDinterval(mcmc(s))
  return(m["var1","lower"])
}
logmeanexp <- function(x){
  x.num <- as.numeric(x)
  xstar = max(x.num)
  return(xstar + log(mean(exp(x.num - xstar))))
}

compute_r2 <- function(df,v1, v2, sigfigs = 3){
  return(round(cor(df[[v1]], df[[v2]])^2, sigfigs))
}

compute_mse <- function(df, v1, v2, sigfigs = 3){
  return(round(mean( (df[[v1]]-df[[v2]])^2), sigfigs))
}

project.path <- "../"
options("scipen"=10) 
```

<!-- 
possible titles:
- Communicating generalizations
- The language of generalization
- The language of generalization: Probability, vagueness, and interaction
-->

# Introduction

Learning that an object tends to have a property, an entity tends to exhibit a behavior, or a cause tends to produce an effect can be crucial to thrive in our open-ended world.
Yet such knowledge can be difficult to acquire: Relevant observations may be costly (e.g., learning that a plant is poisonous) or rare (e.g., understanding that lightning strikes tall objects).
It thus is important that language allow us to communicate such *generalizations* to each other.
By sharing generalizations, we flourish collectively without individually needing to taste potentially-poisonous plants or personally witness many lightning strikes.
Furthermore, communicating generalizations from one generation to the next may be a foundational mechanism behind the faithful transmission of knowledge necessary for cumulative cultural evolution [@Tomasello1999; @Henrich2015].

<!-- Generalizations can be acquired from the world by drawing inductive inferences from observations [@Baldwin1993; @Xu2008; @Dewar2010]. \ndg{cite older stuff. eg hume} -->
<!-- Cooperative teachers can facilitate inductive learning by supplying helpful observations to the learner [@Csibra2009; @Tomasello1999; @Butler2012]. -->
<!-- At around a child's first birthday, however, the learner gains access to a new, human-unique source of data: language. -->
<!-- Minimally, language fosters generalizations by bringing attention to objects being of the same kind through the use of labeling [e.g., "This is a dog"; @Gelman1986; @Gopnik2000; @Graham2004; @Markman1989; @Schulz2008]. -->
<!-- Objects have many properties though; when speakers attach properties to kind-labels (e.g., "Dogs bark"), listeners hear a generalization. -->
<!-- \ndg{this par doesn't really do much} -->

Generalizations in language are often studied via *generic language* (or, generics): simple statements that convey generalizations about categories [e.g., "Dogs are friendly"; @Carlson1977; @Cohen1999; @Leslie2007; @Nickel2008].
Generics are ubiquitous in everyday conversation as well as in child-directed speech [@Gelman2008]. 
In contrast to *particular statements* (e.g., "Rufus is friendly") which describe individual entities or situations and can be verified through observation, generic statements refer to inherently unobservable categories (e.g., the category of \textsc{dogs}) and convey information that extends beyond the present context, a fact which even 2-and-a-half-year-olds appreciate [@Cimpian2008].
<!-- Additionally, generics are the primary way by which speakers discuss social categories, and thus are key to propagating stereotypes [@Gelman2004; @Rhodes2012; @Leslie2015] and impacting motivation [@Cimpian2010]. -->
Generics are just one case of generalization in language, however; events can be described in generalizations (*habitual language*; e.g., "Mary swims after work.") as well as causal relationships (*causal language*; e.g., "Staring at the sun makes you go blind.").
It is believed that every language can express generalizations [@Behrens2005; @Carlson1995] and that such language is central to the growth of conceptual knowledge [@Gelman2004].

<!-- Generic statements ascribing properties to categories are just one kind of generalization found in language.  -->
<!-- Individual entities can also be talked about in generalization, using so-called  -->
<!-- <!-- Recent evidence suggests talking about the other in generalization (using so-called *generic You*) can help make meaning from difficult situations   [e.g., "You never know what will happen on a blind date."; @Orvell2017]. -->
<!-- Communicating about causality can also take the form of communicating a generalization about a causal event  -->
<!-- What these statements have in common is that they convey generalizations, or exhibit *genericity* [@Carlson1977; @Carlson1995]. -->

Despite their ubiquity and relative simplicity, there have been few formal models of generic language and none that have made precise, quantitative predictions to match human intuitions. 
The major challenge in formalizing the language of generalizations is in determining what makes such statements true or false.
When we hear "Dogs bark", it sounds like a universal statement, like *All* dogs bark.
But the property doesn't apply to all dogs: Basenji dogs are barkless^[
http://strangesounds.org/2014/09/basenji-dogs-basenjis-dont-bark-howl.html
], for example.
The generalization could signal something weaker like *Most* dogs bark.
<!-- Indeed, many generalizations seem to have the quantificational force of *most*. -->
However, only a tiny fraction of moquitos carry malaria, while the statement "Mosquitos carry malaria" is intuitively true.
<!-- *most mosquitos* do not carry malaria, while the statement "Mosquitos carry malaria" is intuitively true. -->
The only quantifier that would permit such a statement would be *some* (e.g., "Some mosquitos carry malaria").
But weakening the generalization to *some* goes too far: Half of robins are female, but saying "Robins are female" seems odd.
That same half of robins (i.e., the females), however, also lay eggs, and "Robins lay eggs" is again intuively true.

These observations have led to proposals that the prevalence or probability of a member of the category having the property (e.g., $P(x \text{ barks} \mid x \text{ is a dog})$) is only tangentially related to generic meaning. 
*Conceptual accounts* of generics emphasize the structure of "generic knowledge" [@Prasada2000], and view generic utterances as the way of expressing special, mental relationships between kinds and properties [@Leslie2008; @Prasada2012]. 
From this perspective, the statement "Bishops move diagonally" is true not because bishops tend to move diagonally, but rather because those are the rules of chess.
With this in mind, @Leslie2007 proposes *cognitively primitive* principles that dictate how generics become true or false.
*Characteristic* features are one such principle; "Robins lay eggs" is thus true because it is the mode of reproduction, and reproduction is a characteristic dimension for animal kinds.
*Horrifying* or *striking* information also leads to true generics, which is why "Mosquitos carry malaria" is true despite its low prevalence.

Yet others have observed that generics are not limited to conveying rich, conceptual relations [@Nickel2008; @Nickel2016; @Sterken2015]. 
<!-- Arbitrary relationships are also expressed with generics:  -->
"Barns are red" is true because that is how barn-owners tend to paint their barns; if barn-owners decided to switch to blue, the things they would be painting would still be considered barns.
"Ravens are bigger than toasters" because of the size of ravens and the size of toasters, not because of a special relationship between ravens and toasters.
*Conceptual accounts* must then be supplemented with the proviso that, in certain situations, the statistics of the world do influence human intuitions about generics.

Generic statements have diverse theoretical underpinnings, and the conceptual view highlights this diversity [e.g., @Prasada2013].
But insofar as generics are a single class in language---indeed, insofar as they are a subclass of generalization in language---there should be a general, unified principle that explains what makes the statements true or false.
The same logic should apply to all generalizations in language, including habitual and causal language.
This is the goal of this paper.

<!-- It is thus argued that there are qualitatively different kinds of generic statements (i.e., "striking", "characteristic", "statistical"), each with their own conditions on what makes them true. -->
<!-- This diversity is interesting, but what is common across generics? -->
<!-- Is there some core meaning to *genericity*, a general unified principle that explains what makes generic statements true or false? -->
<!-- Finally, formalizing such a theory is crucial to make explicit the relevant assumptions of the psychological or linguistic hypothesis meant to match human intuitions; this, in turn, makes the theory more testable.  -->

In this paper, we take a critical look at the notion that prevalence---or its more general formulation, probability---cannot serve as the core of a semantic theory of generalization in language (including generics).
We strive for a probability-based semantics for its generality: Probability is the universal currency of belief and serves as a useful representation for human generalization from observations [@Shepard1987; @Tenenbaum2001].
Given that certain probabilities are the *result* of inductive generalization, it would make sense that these are the object of communication for generalizations. 
We formalize such a semantics in a general cognitive modeling framework, and thus incorporate generalizations in language into a larger class of models designed to formalize concepts and common-sense reasoning [@Goodmanconcepts].
Formalizing such a theory is scientifically useful as well to make explicit the assumptions the theory, which in turn, elucidates the conditions under which the theory may be meaningfully tested.

We take the primary desideratum for a theory of the language of generalization to be adult human intuitions into the truth or felicity of statements conveying generalization, which we measure empirically. 
To test the hypothesis that probability can serve as the core semantic theory, we build a computational model of endorsement around this semantic theory and compare our model to human judgments.
Though concepts such as *striking properties* are not explicitly formalized in our model, there is a natural avenue for future integration as our theory is couched in a general cognitive modeling framework.

Specifically, we formalize an account where the core meaning of a generalization in language is a *simple but underspecfied* function of probability and where context helps resolve the underspecification.
In its simplest form, context takes the shape of interlocutors' shared beliefs about the relevant properties or categories under discussion [cf., @Grice1975; @Clark1996; @Levinson2000].
We then ask what a rational speaker---who can only either endorse the generalization or stay silent---would say in different contexts; this is our formal model of endorsement.
We show how the decision to endorse depends in systematic ways on the shared prior knowledge of the interlocutors. \mht{"interlocutors" may be confusing here...}
We test this theory in three distinct case studies---generalizations about categories, events, and causes---by both measuring and manipulating prior knowledge, showing that it is strongly and causally related to the endorsement of generalizations in the ways predicted by our model.
<!-- We show that in each domain, our general theoretical framework explains endorsements about generalizations conveyed in language, while simpler models fall short.  -->

<!-- We think prevalence may have been cast aside too quickly for two main reasons: (i) in previous empirical work, prevalence has been rather narrowly defined, as frequency, where in fact it can be more generally seen as probability (ii) the backround knowledge that adult speakers, whose intuitions are consulted to establish a theory of meaning, may bring to the table has not been explicitly considered or formalized in studies and theoretical work into the language of generalization. -->

<!-- We propose a formal, unified account of generalization in language that can make precise, quantitative predictions about these simple and ubiquituous utterances. -->
<!-- We do so by first considering a much more well-studied form of generalization: generalization from observations.  -->
<!-- The Bayesian probability calculus describes inductive generalizations in terms of probability, and this is has proven to be a good model of human generalization . -->

<!-- We formalize this using a rational model of communication --- the Rational Speech Act (RSA) theory [@Frank2012; @Goodman2013; @Franke2015; @Goodman2016]. -->
<!-- Such models of rational communication describe the inferences listeners can draw from language, and the utterances speakers should say given their state of knowledge and communicative goals.  -->
<!-- Heretofore, however, the fragments of language formalized in such models have been limited to *particular statements* that convey information about a specific entity or situation [e.g., in hyperbole: "My watch cost a million dollars"; @Kao2014, or vagueness: "John is tall"; @Lassiter2013]. -->
<!-- The language of generalization does not make reference to entities in the world or concrete situations, but rather refer to unobservable concepts in the head. -->

<!-- We relate known measurements on factors regarding generic language to each other via our model. -->
<!-- We then go on to manipulate the relevant factors, showing they are causally related via our model to generic endorsement and interpretation.  -->

The paper is organized as follows. 
In Section 1, we describe the Bayesian framework for generalization from observations, serving as the foundation for our semantic theory.
In Section 2, we formalize the hypothesis that the core of meaning of a generalization is *simple but underspecified* and incorporate this *uncertain semantics* into a computational model of endorsement.
<!-- To test our model, we examine endorsements of generalizations in three domains: categories, events, and causes. -->
<!-- Our model makes predictions about how endorsements should vary according to (i) the listener's prior world knowledge and (ii) the speaker's state of knowledge (or, belief distribution) that she seeks to communicate.  -->
In Section 3, we describe experiments that measure both listener prior knowledge and the prevalence of the properties in question, and relate them via our model to the endorsements of generalizations about categories (*generic language*). 
In Section 4, we describe experiments that measure listener prior knowledge and manipulate the frequency of various events to explain endorsements of generalizations about events (*habitual language*).
Finally, in Section 5, we describe experiments that manipulate both the listener's prior knowledge and the prevalence of the properties in question in endorsing generalizations about causes (*causal language*).

Across all of our experimental settings, we find a very strong agreement of our model's predictions to the full spectrum of human elicited endorsements (cumulative $r^2(N) = 0.XX$).
In the final section of the paper, we discuss the generality of our approach and pave a path to unifying the extant psychological, linguistic, and philosophical literatures on generalization in language.
We also show how our endorsement model is a special case of a speaker model in a rational framework of communication. 
This opens the door to understanding generalizations in pragmatic language understanding contexts and building models of learning from language.

<!-- (2) It also allows us to consider sources of gradience in behavior, determining whether or not we should allocate it to *noise* in the observed data or to *uncertainty* in the head of the language user. -->
<!-- *Causal language* (e.g., "The block *makes* it [the machine] go.") faciliates more abstract forms of causal reasoning [@Bonawitz2010]. -->

<!-- \ndg{There haven't been many formal models of generic language, none have made quantitative predictions. Say why this is important. -->
<!-- To build one, we need to firsts say what a generalization is. We use the Bayesian idea that an inductive generalization can be understood in terms if probbilities. -->
<!-- Then we need to say how these probabilities are communicated. Use RSA...} -->

<!-- Rational models of communication have had a lot of recent success in formalizing complex linguistic phenomena ranging from politeness [@Yoon2016] and hyperbole [@Kao2014] to metaphor [@Kao2014] and vagueness [@Lassiter2013; @Qing2014].  -->

<!-- give rise to the productivity of human thought and form of foundations of cultural knowledge. -->
<!-- They allow us to understand the complex world around us and dream up how possible future situations might unfold. -->
<!-- It's been suggested that were psychology to have *general laws* akin to Newtonian laws in physics, the first general law in psychology would be of generalizations [Shepard1987]. -->
<!-- Generalizations are central to human understanding.  -->
<!-- Even before birth, human and non-human animals alike gather statistical information and form generalizations to predict what will come next. -->
<!-- The human learner, perhaps uniquely in the animal kingdom, displays striking sensitivity to the intentional and pedagogical actions of others, leading to distinct and often stronger inferences from data than would otherwise be warranted from random observations. -->
<!-- and about people in general [e.g., "You never know what will happen on a blind date."; @Orvell2017]. -->

<!-- It's plausible that cultural knowledge accumulates across generations through language's ability to talk about generalizations [@Henrich2015]. -->

<!-- Generic statements convey generalizations about categories, but there are other linguistic devices that look different but seem to similary convey generalization. -->
<!-- *Genericity* can be observed when talking about categories (e.g., *Dogs bark*), people in general , specific people (e.g., *Martha swims after work*), or causal events (e.g., *Drinking moonshines makes you go blind*). -->
<!-- Though this diverse array of utterances all seem to convey something beyond the here and now, what exactly is common to each of them is not obvious.  -->
<!-- Without a formal way of describing generalizations in language, we do not yet know in what ways these large swaths of language are similar.  -->

<!-- The language of generalization is exceedingly difficult to formalize, however. -->
<!-- The cognitive act of generalization is naturally formalized in Bayesian cognitive models [@Tenenbaum2011]. -->
<!-- Bayesian models showcase the role of flexibly structured representations [@Kemp2007; @Kemp2008] and assumptions learners can make about the data-generating process [e.g., learning from pedagogical actions; @Shafto2012]. -->
<!-- Learning from language, however, is more challenging to formalize because of the inherent context-sensitivity of language [@Grice1975; @Clark1996; @Levinson2000]. -->

<!-- Exceptions could be made for properties that only apply to one gender (like being female), but *Robins lay eggs* is intuitively true despite the property also being only present in females. -->
<!-- Generics roughly convey that the properties is widespread in the category (e.g., *Most dogs are friendly*), however they don't always imply the same degree of prevalence. -->
<!-- *Birds are female* sounds strange even though half of birds of female.  -->

<!-- ALTERNATIVE INTRO -->

<!-- Yet, generalizations can be hard to acquire. -->
<!-- Consider what it would take to learn that under certain conditions (e.g., when eaten raw) a plant is poisonous, while under other conditions (e.g., when processed appropriately) the plant is nutritious.  -->
<!-- This is a common ecological problem and historical examples suggest that humans are not particularly good at learning such generalizations [e.g., the Australian explorers case; @Henrich2015]. -->
<!-- In this domain, the relationship to be learned (i.e., between different ways of food processing and the poisonousness of the plant) is non-obvious and trial-and-error learning is costly (negative examples entail death). -->

<!-- Crucially, once acquired, such hard-won generalizations can be communicated among humans, something which other animals struggle to accomplish [cite tomasello apes something]. -->
<!-- Generalizable knowledge can be inferred through demonstration, or in pedagogical contexts when ostensive cues are present [@Csibra].  -->
<!-- By and large, however, language provides the most flexible and generative way to communicate generalizations.  -->

<!-- A generalization provides a relation between a category and a property: It tells us to predict that property of future instances of the category [@HumeTHN]. -->
<!-- Generalization can be described as a probability and inductive inference as a probabilistic model [@Tenenbaum2011]. -->
<!-- From a very young age, children make strong inductive inferences from their experience in a way consistent with the Bayesian probability calculus [@Dewar2010; @Gelman1986; @Markman1989]. -->

<!-- If generalizations can be described by predictive probabilities, it would stand to reason that the most direct way to communicate a generalization would be to convey probability.  -->
<!-- Ironically, humans are very bad at describing and interpreting probabilities explicitly [@Tversky1974]. -->
<!-- Rather than be a deficiency, however, the difficulty we have in understanding numerical descriptions of probability might reflect the flexible adaptivity of natural language to make vague or ambiguous that which shared context can be relied upon to guide interpretation [@Piantadosi2012]. -->
<!-- If language has evolved in a way to make efficient the kinds of inductive inferences that are central to human cognition, we would expect the language of generalizations to be simple [c.f., @Leslie2008]. -->

<!-- Indeed, generalizations in language (e.g., *Dogs are friendly*, *Mary smokes cigarettes*, *Yeast makes bread rise*) are among the simplest forms of compositional expressions.  -->
<!-- Generalizations do not come in just one syntactic form: "*Dogs* are friendly", "*A library book* should be treated with respect", "*The contemporary foodie* seeks culinary experinces in menu-less restaurants and grandmothers' kitchens" [@Carlson1995; @NickelBlackwell; @sepGenerics]. -->
<!-- Each of these statements has in common the fact that they predicate a property of a category and the properties themselves apply to instances of the category (unlike, e.g., *Dinosaurs are extinct* because being extinct is not a property that can be true of a particular dinosaur).  -->
<!-- In this way, these statements exhibit *genericity*, or put another way, they convey generalizations. -->

<!-- In this paper, we formalize the language of generalizations with the interest is deciding what makes a generalization in language *true or false*. -->
<!-- How speakers convey generalizations and how listeners interpret them. -->

<!-- We examine three case studies of generalizations in language: Generalizations about agents, objects, and events.  -->
<!-- In Section 1, we introduce an information-theoric communicative theory of how generalizations are conveyed in language and flesh out the model's predictions alongside alternative models. -->
<!-- In order to do so, we provide a unified mathematical framework for dealing with agents, objects, and events, casting each as a category over which generalization can occur. -->
<!-- Our model predicts the endorsements of generalizations should vary with the (A) the probability of an instances having the property and (B) the prior distribution over the property probability (e.g., how likely instances of other categories are to have the property). -->
<!-- In Section 2, we introduce our empirical paradigm which we use to measure endorsement and interpretation of linguisitic generalizations, examining generalizations about *agents* in what is known as *habitual language* (e.g., "Mary smokes cigarettes.").  -->
<!-- In Section 3, we extend our paradigm to generalizations about *events* (i.e., *causal language* e.g., "Yeast makes bread rise.") and experimentally manipulate participants' background knowledge, showing that prior beliefs about probabilities are causally related to endorsement and interpretation of generalizations. -->
<!-- In Section 4, we use our paradigm to address long-standing theoretical and empirical puzzles regarding generalizations about *objects*, so-called *generic language*.  -->
<!-- We show our how model explains three known psychological puzzles surrounding generic language.  -->
<!-- In all case studies, we provide formal comparison to simpler models and in all cases, find that only our communicative theory explains the patterns of empirical data. -->
<!-- We conclude with a general discussion about the relationship of our theory to extant theories in the psychology, linguistics, and philosophy of genericity, and highlight the implications for modeling language acquisition, conceptual development, and other avenues of research that this framework opens. -->

<!-- the truth conditions of generalizations of events (so called *habitual language*) by manipulating the target propensity (A) while surveying a wide variety of events (which should induce variability in B). -->
<!-- In Section 3, we manipulate both (A) and (B) while testing the model on generalizations about causes (*causal language*). -->
<!-- If our theory is truly a theory about communicating generalizations, it should be able to explain the known psychological puzzles surrounding *generic language* (generalizations about categories).  -->

<!-- In stark constrast, a computational understanding of communicating generalizations is not well understood at all.  -->

<!-- Gathering data and observing evidence, however, is costly.  -->
<!-- The human problem of induction even displays a complex sensitivity to not only the type of data, but how the data was generated [@Gweon2010; @Shafto2012]. -->
<!-- A learner cannot form all of her abstract knowledge through direct experience, however. -->
<!-- Life is only so long, and you can only be in one place at a time. -->

<!-- In this paper, we introduce a formal model for understanding the language of generalizations. -->
<!-- <!-- The fact that generalizations in language seem to convey something about the propensity but that there seem to be as many interpretations of generalized statements as there are generalizations have confounded formal models aimed to capture their meaning. -->
<!-- We take as a starting point the idea that the core meaning of a linguistic generalization is simple (i.e., it conveys something about the propensity), but underspecified (i.e., there is uncertainty in the meaning). -->
<!-- We extend a rational model of communication --- the Rational Speech Act theory --- to be able to understand general statements that extend beyond the here-and-now. -->
<!-- We provide a number of stringest tests of this theory to explain generalizations about categories, events, and causal forces.  -->

<!-- How can generalizations in language have such flexible truth conditions while simultaneously carrying intuitive and, at times, very strong implications? -->
<!-- This question has been investigated by psychologists, linguists, and philosophers since @Carlson1977. -->
<!-- Here, we provide a unified formal theory of generalizations conveyed in language. -->
<!-- We test theory on the preeminent case study of linguistic generalizations: generic language, or generalizations about categories. -->
<!--  -->
<!-- To be precise, we develop a computational model that describes pragmatic reasoning about the propensity required to assert the generaization.   -->
<!-- We find that this formalism resolves known philosophical and empirical puzzles. -->
<!--We also provide the first set of experimental results about linguistic generalizations about events and causal forces, so called *habitual* and *causal statements*.-->

<!-- and find extreme flexibility in usage, which our model predicts.  -->

<!-- In Section 2, we conduct a set of experiments to test the *truth conditions* of familiar generalizations about categories (generic statements), and find our model predicts key patterns in human judgments -->
<!-- In Section 3, we conduct a set of experiments to test *interpretations* of novel generic sentences, again finding our model predicts human judgments with high quantitative accuracy. -->
<!-- In Section 4, we begin to test the generality of the theory, examining the *truth conditions* of generalizations about events. -->
<!-- In Section 5, we further probe our theory's generality, investing the *truth conditions* and *interpretations* of novel generalizations about causal forces.  -->
<!-- In all cases, we provide formal comparison to simpler models and in all cases, find that only our communicative theory with an underspecified meaning function explains the pattern of human data. -->
<!-- We also find that the prior beliefs used in our model reflect conceptual structure, and this provides an understanding of the conceptual implications of generics. -->
<!-- Finally, we investigate the underlying semantic scale in more detail, and find that the core meaning of generalizations in language depends in a crucial way on subjective beliefs, not mere frequency. -->

<!--
Most would agree that \emph{Swans are white}, but certainly not every swan is.
This type of utterance conveys a generalization about a category (i.e. \textsc{swans}) and is known as a generic utterance [@Carlson1977; @Leslie2008].
Communicating generically about categories is useful because categories themselves are unobservable [@Markman1989].
Knowledge about categories, while central to human reasoning, is tricky to acquire because categories themselves are unobservable [Markman1989}.
It is believed that every language can express generic meaning [@Behrens2005; @Carlson1995], and that generics are essential to the growth of conceptual knowledge [@Gelman2004] and how kinds are represented in the mind [@Leslie2008].
Generic language is ubiquitous in everyday conversation as well as in child-directed speech [@Gelman2008], and children as young as two or three understand that generics refer to categories and support generalization [@Cimpian2008].
and though English and many other languages do not possess an unambiguous form devoted to generic meaning [@Behrens2000, @AlMalki2014].
Additionally, generics are the primary way by which speakers discuss social categories, making them key to propagating stereotypes [@GelmanEtAl2004; @Rhodes2012; @Leslie2015] and impacting motivation [@Cimpian2010motivation].
Despite their psychological centrality, apparent simplicity, and ubiquity, a formal account of generic meaning remains elusive.

How can generics have such flexible truth conditions while simultaneously carrying strong implications?
In this paper, we explore the idea that the core meaning of a generic statement is simple, but underspecified, and that general principles of communication may be used to resolve precise meaning in context. 
In particular, we develop a mathematical model that describes pragmatic reasoning about the degree of prevalence required to assert the generic.  
We find that this formalism resolves the philosophical and empirical puzzles.
-->

# Computational Framework

Generalizations are used to make predictions about properties of instances that an agent has yet to experience [@HumeTHN].
People readily predict that a new dog will have fur, drinking a new cup of coffee will cause jitters, and a new day will find certain people typing.
In each case, we assign a specific exemplar $x$ to a category $k$, and make a prediction that it will have property $f$.
This prediction can be described by a conditional probability: $h = P(x \in f \mid x \in k)$, the probability that $x$ will have $f$ given that it is in $k$, which we will refer to as the **feature-probability**.
The targets of our predictions can vary widely: They may be objects (a dog), events (a coffee drink), or more fine grained types (a person on a given day).
The properties correspondingly also vary widely (e.g., having fur, causing jitters, typing). 
Yet the mathematical description of the inductive inference is always given by the conditional feature-probability $h$.

How does one come to learn $h$, the uncertainty that underlies an inductive inference about a member of a category having a property?
One way is through observation: One can observe instances of the category $k$ and note how many of them have $f$.
But the inferences derived from observation will depend upon on one's background knowledge about the property [@Gelman2003; @Keil1992].
For example, if you observe an instance of a novel creature category that \emph{has two legs}, you may infer that all of that category has two legs.
If in addition you observe that it has broken wings, the corresponding prediction of another instance having broken wings is not nearly as strong [@Nisbett1983].
This kind of knowledge can be represented formally as a probability distribution over the feature-probabilities $h$: $P(h)$ [@Kemp2008], and observational data $d$ can update one's beliefs via Bayes' Rule: $P(h \mid d) \propto P(d \mid h) \cdot P(h)$
Here, $P(d \mid h)$ is the observer's generative model of the observed data. 
It is conditional on $h$ because the probability of various direct observations depends upon the feature-probability that the observer assumes.

Generalization from observational evidence is limited, however, in its capacity to describe how the mind builds complex theories of the world or how culture accumulates across generations.
Relying upon observations would be particularly problematic, for example, to learn about properties that are costly to observe (e.g., *staring at the sun makes you go blind*) or events that are statistically rare (e.g., *lightning strikes tall objects*).
How do we acquire these generalizations?
For abstract, generalizable knowledge to remain in a culture, it must be faithfully transmitted between individuals and across generations [@Tomasello1999], and language provides simple ways to communicate generalizations.

<!-- The human capacity to generalize from observations, deriving $h$ from example $x$s, has been studied extensively in cognitive and developmental psychology [@Nisbett1983; @Baldwin1993; @Heit2000]. -->
<!-- From a very young age, children draw strong inferences about nonobvious object properties from just a few examples [@Baldwin1993; @Gopnik2000;  @XuTenenbaum2007]. -->
<!-- Such inductive generalization can be described by the Bayesian probability calculus: An observer's initial beliefs or state of knowledge---called a prior belief distribution---is updated through experience (i.e., after making observations) and becomes a posterior belief distribution (what they belief *a posteriori*).  -->
<!-- This posterior distribution becomes the observer's prior for the next observation she makes: it thus informs her predictions about unobserved instances. -->

<!-- As a concrete example, consider a learner investigating the feature \textsc{barks} ($f$) for the category \textsc{dog} ($k$). -->
<!-- $h$ describes the learner's subjective probability that the next dog she encounters will bark^[At the moment, we assume the feature \textsc{barks} is an all-or-none Boolean property. Later, we will see how the presence of this feature itself can be thought of as a generalization over events of barking.], or more abstractly, her belief about the barking capacity of dogs. -->
<!-- That is, $h = P(x \text{ barks} \mid x \text{ is a dog})$. -->
<!-- The learner does not know $h$, but has some idea about what $h$ could be; this is represented by $P(h)$, the learner's prior beliefs about how likely it is that $x$ will bark given that $x$ is a dog. -->

>>>>

```{r echo = F}
library(langcog)
library(tidyverse)
library(ggthemes)
library(jsonlite)
theme_set(theme_few())
project.path <- "../"
data.path <- "data/generics/endorsement/"
substrRight <- function(x, n){
  substr(x, nchar(x)-n+1, nchar(x))
}
```

# Appendix: Bayesian Data Analysis

## Case Study 1: Generic language

### Modeling feature-probability priors

In Case Study 1 (Generic Language), we elicited the prevalence of features for many different categories. 
These can be thought of as *samples* from the feature-probability prior distribution. 
Formally, we assume the prior data (analyzed independently for each property) was generated from one of two distributions: a distribution corresponding to those kinds with a stable causal mechanism that *could* give rise to the property ($\mathcal{D}_{stable}$) and a "transient cause" distribution corresponding to those kinds without a stable mechanism ($\mathcal{D}_{transient}$).
The "transient" distribution intuitively corresponds to accidental causes of the feature (e.g., a lion, who through some genetic mutation, reproduces by laying eggs).
We model this distribution as a Beta distribuition that heavily favors probabilities near 0: $\text{Beta}(\gamma = 0.01, \delta = 100)$.^[
  Note that we use the noncanonical mean $\gamma$ and concentration $\xi$ (or, inverse-variance) parameterization of the Beta distribution rather than the canonical shape (or pseudocount) parameterization for ease of posterior inference. The shape parameterization can be recovered using: $\alpha = \gamma \cdot \xi; \beta = (1 - \gamma) \cdot \xi$.
]
The "stable" distribution is modeled as a Beta distribution with unknown parameters $\text{Beta}(\gamma, \xi)$.^[
Because the Beta distribution is not defined at the points 0 and 1, we add $\epsilon$ to the 0 responses and round 1 to 0.99.
Adjusting 1 to $1- \epsilon$ leads to improper inferences for this model, as $1 - \epsilon$ is only likely under a highly right-skewed distribution; treating 1 as $1- \epsilon$ disproportionately influences the shape of $\mathcal{D}_{stable}$ relative to other responses.
This problem does not appear for 0 being adjusted to $\epsilon$ because the "transient" distribution already expects such low values.
Similar results can be obtained by rounding 0 to 0.01. 
Alternatively, the "transient" distribution could be defined as a Delta distribution at 0, and 0 responses could remain in their raw form.
]
Finally, we assume that these two components combine with mixture weighting $\phi$ such that the data we observe is $$P(d) = \phi\cdot \text{Beta} (d \mid \gamma, \xi) + (1 -  \phi) \cdot \text{Beta}(d \mid \gamma = 0.01, \xi = 100) $$.
<!-- The unstructured alternative hypothesis assumes the data was generated by a single component in the prior distribution with unknown parameteres $P(d) = \text{Beta} (d \mid \gamma, \xi)$. -->
We put the following priors over the latent parameters of the model:
\begin{eqnarray}
\phi_i & \sim & \text{Uniform}(0, 1) \\
\gamma_i & \sim & \text{Uniform}(0, 1) \\
\xi_i & \sim & \text{Uniform}(0, 100)
\end{eqnarray}
where $i$ ranges over the different properties (e.g., *lays eggs*, *carries malaria*).
